Classless Java

### 2015-11-15 paper/sections/Introduction.tex
added references

Java 8 introduced \emph{default methods}, allowing interfaces to have
method implementations. The main motivation behind the introduction of
default methods in Java 8 is \emph{interface evolution}. That is, to
allow interfaces to be extended over time, while preserving backwards
compatibility. It soon became clear that default methods could also be
used to emulate something similar to \emph{traits}~\cite{}. The
original notion of traits by Scharli et al. prescribes, among other
things, that: 1) a trait provides a set of methods that implement
behaviour; and 2) a trait does not specify any state variables, so the
methods provided by traits do not access state variables
directly. Java 8 interfaces follow similar principles too. Indeed, a
detailed description of how to emulate trait-oriented programming in
Java 8 can be found in the work by Bono et al.~\cite{}.  The Java 8
team designing default methods, was also fully aware of that secondary
use of interfaces, but it was not their objective to model traits:
``The key goal of adding default methods to Java was "interface
evolution", not "poor man's traits"''~\cite{mailing list}. As a result
they were happy to support the secondary use of interfaces with
default methods as long as it did not make the implementation and
language more complex.

@@

\item {\em Useful, general purpose methods, require special care in
  the presence of subtyping.} Methods such as $clone$, or
  \emph{fluent} setters~\cite{}, not only require access to the
  internal state of an object, but they also require their types to be
  refined in subtypes.

@@

To evaluate the usefulness of object interfaces, we illustrate 3
applications and case studies. The first application is a simple 
solution to the Expression Problem~\cite{}, supporting independent 
extensibility~\cite{}, and without boilerplate code. The second
application is to show how embedded DSLs using fluent interfaces~\cite{} 
can be easily defined using object interfaces. Finally, the last
application is a larger case study for a simple Maze game implemented with 
multiple inheritance. For the last application we show that there is a
significant reduction in the numbers of lines of code when compared 
to an existing implementation~\cite{} using plain Java 8. 
Noteworthy, is the fact that all applications are implemented 
without defining a single class.

>>>>

Java 8 introduced \emph{default methods}, allowing interfaces to have
method implementations. The main motivation behind the introduction of
default methods in Java 8 is \emph{interface evolution}. That is, to
allow interfaces to be extended over time, while preserving backwards
compatibility. It soon became clear that default methods could also be
used to emulate something similar to \emph{traits}~\cite{scharli03traits}. The
original notion of traits by Scharli et al. prescribes, among other
things, that: 1) a trait provides a set of methods that implement
behaviour; and 2) a trait does not specify any state variables, so the
methods provided by traits do not access state variables
directly. Java 8 interfaces follow similar principles too. Indeed, a
detailed description of how to emulate trait-oriented programming in
Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
team designing default methods, was also fully aware of that secondary
use of interfaces, but it was not their objective to model traits:
``The key goal of adding default methods to Java was "interface
evolution", not "poor man's traits"''~\cite{goetz13default}. As a result
they were happy to support the secondary use of interfaces with
default methods as long as it did not make the implementation and
language more complex.

@@

\item {\em Useful, general purpose methods, require special care in
  the presence of subtyping.} Methods such as $clone$, or
  \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
  internal state of an object, but they also require their types to be
  refined in subtypes.

@@

To evaluate the usefulness of object interfaces, we illustrate 3
applications and case studies. The first application is a simple 
solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
application is to show how embedded DSLs using fluent interfaces~\cite{} 
can be easily defined using object interfaces. Finally, the last
application is a larger case study for a simple Maze game implemented with 
multiple inheritance. For the last application we show that there is a
significant reduction in the numbers of lines of code when compared 
to an existing implementation~\cite{bono14} using plain Java 8. 
Noteworthy, is the fact that all applications are implemented 
without defining a single class.
